---
title: "Assignment 3 Task 1"
author: "Jessica Jagdeo"
date: "3/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
```

####Task 1: SB creek water chemistry, multivariate exploration & analysis

You have a number of tools in your multivariate analysis & exploration toolbelt, like:

- Multiple linear regression
- Correlation matrices
- Principal components analysis
- K-means clustering
- Hierarchical clustering

For Task 1, you should explore, analyze and visualize a multivariate dataset using at least 3 of the methods listed above (more if you feel inspired). Prepare your analysis and findings in a single knitted HTML that can be featured as a project to highlight your data analysis and coding skills (e.g. on yourpersonal website). The analysis should provide a short introduction, clear subsections and descriptions, and insights/summaries of relationships between variables and between groups.

You can decide which variables and observations you want to explore/include. For example: You might narrow the data to just a few sites of interest, or only include complete cases (with no NAs) for cluster analyses. Hint: if you’re doing hierarchical clustering by site , you’ll want to make a data frame that has
a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance.

#### 1. Load packages and read in data
```{r}

library(tidyverse)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)

```

```{r}

sb_creek <- read_csv("sbc_lter_registered_stream_chemistry.csv")

# NH4: ammonium
# NO3: nitrate
# PO4: phosphorus
# TDN: total dissolved nitrogen
# TDP: total dissolved phosphorus
# TPC: total particulate carbon
# TPN: total particulate nitrogen
# TPP: total particulate phosphorus
# TSS: total suspended solids
# Spec_Cond: specific conductivity

```

#### 2. Examine and remove NAs
```{r}

sb_creek_na <- na_if(sb_creek, -999.0)

summary(sb_creek_na)

gg_miss_var(sb_creek_na)

# Remove the columns in which most observations are NAs and remove NAs from rest of the columns

sb_creek_clean <- sb_creek_na %>% 
  select(-tpc_uM,
         -tpn_uM,
         -tpp_uM,
         -tss_mgperLiter) %>% 
  drop_na()

# Reduced dataset from 19,000 to 12,000 observations

skim(sb_creek_clean)

```

#### 3. Principle components analysis
```{r}

sb_creek_pca <- sb_creek_clean %>% 
  select(-site_code,
         -timestamp_local)

sb_creek_chem_pca <- prcomp(sb_creek_pca, scale = TRUE)

summary(sb_creek_chem_pca)

# Plot PCA results to visualize associations

sb_creek_biplot <- autoplot(sb_creek_chem_pca, 
                      colour = "NA",
                      loadings.colour = "blue",
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
                      theme_classic() 
  

sb_creek_biplot

# Biplot captures about 65% of variance

```

**Figure 1. Principal components analysis biplot showing the associations between chemistry samples taken from Santa Barbara Creek.** Variables depicted by this biplot include ammonium, nitrate, phosphorus, total dissolved nitrogen, total dissolved phosphorus, and specific conductivity. Data source: Melack, J.

To summarize key observations from Figure 1:

1) There is a strong, positive correlation between phosphorus and total dissolved phosphorus.

2) There is a positive correlation between nitrogen and total dissolved nitrogen (not as strong as the relationship in 1).

3) There is no correlation between specific conductivity and phosphorus/total dissolved phosphorus.

#### 4. Cluster analysis
```{r}

# Asking R: How many clusters do you think there should be for this dataset?

# number_est <- NbClust(sb_creek_clean[3:8], 
#                       min.nc = 2,
#                       max.nc = 5,
#                       method = "kmeans")

# Do kmeans:
sb_creek_km <- kmeans(sb_creek_clean[3:8], 3)

# Bind the cluster number together with the original data:
sb_creek_cl <- data.frame(sb_creek_clean, cluster_no = factor(sb_creek_km$cluster))

# Plot my different clusters:
ggplot(data = sb_creek_cl) +
  geom_point(aes(x = no3_uM, y = nh4_uM, color = cluster_no)) +
  theme_classic()

```

**Figure 2. Cluster analysis via k means to categorize SB Creek chemistry samples into three clusters.** Data source: Melack, J.

#### 5. Multiple linear regression
```{r}

sb_creek_lm <- lm(spec_cond_uSpercm ~ nh4_uM + no3_uM + po4_uM + tdn_uM + tdp_uM, data = sb_creek_clean)

summary(sb_creek_lm)

plot(sb_creek_lm)

# This is a really bad model.... These chemistry values do not significantly predict specific conductivity

# Probably should not be running multiple linear regression for this data set!

```

